---
title: "Using the rerdapXtracto routines"
author: "Roy Mendelssohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rerdapXtracto}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r initialize, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(rerddapXtracto)
```

## Introduction 

`rerdapXtracto` is an <span style="color:blue">R</span> package developed to subset and extract satellite and other oceanographic related data from any <span style="color:blue">ERDDAP</span> server using the <span style="color:blue">R</span> package `rerddap` developed by Scott Chamberlain and the wonderful people at [rOpenSci](https://ropensci.org). <span style="color:blue">ERDDAP</span> is a simple to use yet powerful web data service developed by Bob Simons. `rerdapXtracto` extends the `rerddap` package by being able to extract data for a moving point in time along a user-supplied set of longitude, latitude and time points;  and also extracting data within a polygon (through time).  `rerdapXtracto` extends the functionality of the <span style="color:blue">R</span> package `xtractomatic` by being able to work with (hopefully) most gridded datasets available from any <span style="color:blue">ERDDAP</span> server. The disadavantage compared to `xtractomatic` is that the user has to do more work to obtain information about the dataset to be accessed, while in `xtractomatic` that information is built in.

### The Main xtractomatic functions

There are three main data extraction functions in the `rerddapXtracto` package: 

- `rxtracto <- function(dataInfo, parameter = NULL, xcoord=NULL, ycoord=NULL, zcoord = NULL, tcoord=NULL, xlen=0., ylen=0., xName='longitude', yName='latitude', zName='altitude', tName='time', urlbase='http://upwell.pfeg.noaa.gov/erddap', verbose=FALSE)`

- `rxtracto_3D <- function(dataInfo, parameter = NULL, xcoord=NULL, ycoord=NULL, zcoord = NULL, tcoord=NULL, xName='longitude', yName='latitude', zName='altitude', tName='time', urlbase='http://upwell.pfeg.noaa.gov/erddap', verbose=FALSE)`

- `rxtractogon <- function(dataInfo, parameter, xcoord=NULL, ycoord=NULL, zcoord = NULL, tcoord=NULL, xName='longitude', yName='latitude', zName='altitude', tName='time', urlbase='http://upwell.pfeg.noaa.gov/erddap', verbose=FALSE)`

The functions are similar to but not identical to the functions in `xtractomatic`, the main differences being having to obtain information about the dataset first using `rerddap`, and possibly having to give the names of the coordinate variables, as these can't be assumed  (for example the zcoord could be in sigma coordinates).  More specifically:

- dataInfo: the return from an rerddap "info" call to an <span style="color:blue">ERDDAP</span> server
- parameter: character strig containing the name of the parameter to extract
- xName: character string with name of the xcoord in the <span style="color:blue">ERDDAP</span> dataset (default "longitude")
- yName: character string with name of the ycoord in the <span style="color:blue">ERDDAP</span> dataset (default "latitude")
- zName: character string with name of the zcoord in the <span style="color:blue">ERDDAP</span> dataset (default "altitude")
- tName: character string with name of the tcoord in the <span style="color:blue">ERDDAP</span> dataset (default "time")
- urlbase: base URL of the <span style="color:blue">ERDDAP</span> server being accessed - default "http://upwell.pfeg.noaa.gov/erddap"
- verbose: logical variable (default FALSE) if the the URL request should be verbose

### Time has come today

With all due respect to the [Chambers Brothers](https://www.youtube.com/watch?v=uHfB63ln1Ig), since any <span style="color:blue">ERDDAP</span> served gridded data can be accessed, a lot more care must be used with the values pf "time" passed to `rerddapXtratco` as compared with `xtractomatic`.  This is because datasets can have time increments of less than a day, an example of which is [given below](#hourly). <span style="color:blue">ERDDAP</span> maps all times to "Zulu" time, of the form "2016-11-01T00:00:00Z".  The date-time "2016-11-01" gets mapped to "2016-11-01T00:00:00Z".  When giving time limits of finer resolution than a day, be certain to take into account time offsets, and to give the times in the form above.  Times are parsed using the `parse_date()` function of the `parsedate` package, if in doubt you can use that function to see how the times you are passing will be interpreted.


## Setting up 

`rerdapXtracto` uses the `rerddap`, `ncdf4`, `sp` and `parsedate` packages, and these packages (and the packages imported by these packages) must be installed first or `rerdapXtracto` will fail to install.  

```{r install,eval=FALSE}
install.packages("rerddap", dependencies = TRUE)
install.packages("ncdf4",dependencies = TRUE) 
install.packages("sp", dependencies = TRUE)
install.packages("parsedate", dependencies = TRUE)
```

The `rerdapXtracto` package is not available through <span style="color:blue">CRAN</span> at the moment, but it is available from [Github](https://github.com/rmendels/rerdapXtracto)
 
To install the `rerdapXtracto` package from <span style="color:blue">Github</span> ,

```{r installGit,eval=FALSE}
install.packages("devtools")
devtools::install_github("rmendels/rerdapXtracto")
```

Once installed, to use `rerdapXtracto` do

```{r load,eval=FALSE}
library("rerdapXtracto")
```

If the other libraries  (`rerddap`, `ncdf4`,`sp`, and `parsedate`) have been installed they will be found and do not need to be explicitly loaded.   


### Using the R code examples

Besides the `rerdapXtracto` packages, the examples below depend on `akima`, `dplyr`, `ggplot2`, `ggfortify`, `lubridate`, `mapdata`,  `reshape2`,and `xts`.  These can be loaded beforehand (assuming they have been installed):

```{r loadpacks,eval=FALSE}
library("akima")
library("dplyr")
library("ggplot2")
library("ggfortify") 
library("lubridate")
library("mapdata")
library("reshape2")
library("xts")
```

In order that the code snippets can be more stand-alone, the needed libraries are always `required` in the examples.  

It should be emphasized that these other packages are used to manipulate and plot the data in <span style="color:blue">R</span>, other packages could be used as well.  The use of `rerdapXtracto` does not depend on these other packages.

There are also several <span style="color:blue">R</span> functions defined within the document that are used in other code examples.  These include `mapFrame`, and `plotFrame`.



## Getting Started

The first several examples reproduce some of the examples in the `xtractomatic` vignette, hopefully to make clear how the two sets of functions relate. One change is the plots use the `cmocean` colormaps designed by Kristen Thyng (see http://matplotlib.org/cmocean/ and https://github.com/matplotlib/cmocean).  These colormaps were initally developed for Python, but a version of the colormaps is used in the `oce` package by Dan Kelley and Clark Richards and that is what is used here.
 
The first step is to find the necessary information from the <span style="color:blue">ERDDAP</span> server being used.  These include the

- base URL of the <span style="color:blue">ERDDAP</span> server (the default is http://upwell.pfeg.noaa.gov/erddap)
- the dataset ID of the data to be accessed (for example the 1-day composite of chloropyhll from MODIS Aqua has a dataset ID of 'erdMBchla1day' - see http://upwell.pfeg.noaa.gov/erddap/griddap/erdMBchla1day.html)
- the parameter name (in this case "chlorophyll")

The information about the then must be obtained by making a call to the `rerddap` function `info()`:

```{r info}
require(rerddap)
## base URL does not need to given because it is the default one
dataInfo <- info('erdMBchla1day')
dataInfo
```

The `info()` function returns most of the information needed to make a function call in `rerddapXtracto`, to wit the coordinate variables and their limits, as well as the parameters available from the selected dataset.

### An rxtracto example 

In this section we extract data along a trackline found in the <span style="color:red">Marlintag38606</span> dataset, which is the track of a tagged marlin in the Pacific Ocean (courtesy of Dr. Mike Musyl of the Pelagic Research Group LLC), and show some simple plots of the extracted data.  



```{r getMarlinChl}
require(rerddapXtracto)
require(rerddap)

# First we will copy the Marlintag38606 data into a variable 
# called tagData  so that subsequent code will be more generic.  

tagData <- Marlintag38606
xpos<-tagData$lon
ypos<-tagData$lat
tpos<-tagData$date
zpos <- 0.
urlbase <- 'http://upwell.pfeg.noaa.gov/erddap'
swchlInfo <- rerddap::info('erdSWchla8day')
swchl <- rxtracto(swchlInfo, parameter='chlorophyll', xcoord=xpos, ycoord=ypos, tcoord=tpos, zcoord=zpos, xlen=.2, ylen=.2)
```

Note the differences from the `xtractomatic` example.

- the user has to find the datasetID from the given <span style="color:blue">ERDDAP</span> server, in this case 'erdSWchla8day' (from http://upwell.pfeg.noaa.gov/erddap/griddap/erdSWchla8day.html)
- the user has to call `rerddap::info('erdSWchla8day')` to obtain information about that dataset
- the user has to know the parameter name desired for that datasetID, in this case 'chlorophyll'
- the zcoord has to be explictily given,  because 'erdSWchla8day' is a 4-D dataset

### Plotting the results

We plot the track line with the locations colored according to the mean of the satellite <span style="color:blue">chlorophyll</span> around that point. Positions where there was a tag location but no <span style="color:blue">chlorophyll</span> values are also shown.

```{r meantrackPlot, fig.align='center',fig.width=6,fig.height = 3}
require(ggplot2)
require(mapdata)
# First combine the two dataframes (the input and the output) into one, 
# so it will be easy to take into account the locations that didn’t 
# retrieve a value.

alldata <- cbind(tagData, swchl)

# adjust the longitudes to be (-180,180)
alldata$lon <- alldata$lon-360
# Create a variable that shows if chla is missing
alldata$missing <- is.na(alldata$'mean chlorophyll')*1
alldata$mean <- as.array(alldata$'mean chlorophyll')
# set limits of the map
ylim <- c(15, 30)
xlim <- c(-160, -105)
myColor <- colors$chlorophyll
# get outline data for map
w <- map_data("worldHires", ylim = ylim, xlim = xlim)
# plot using ggplot
z <- ggplot(alldata,aes(x=lon,y=lat)) + 
   geom_point(aes(colour=mean, shape=factor(missing)), size=2.) + 
   scale_shape_manual(values=c(19, 1))
z + geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") + 
  theme_bw() + 
  scale_colour_gradientn(colours= myColor,limits=c(0., 0.32),"Chla") + 
  coord_fixed(1.3, xlim = xlim, ylim = ylim) + ggtitle("Mean chla values at marlin tag locations")
```


### Topography data

The second example from the  `xtractomatic`  vignette is accessing topographic data along the Marlin track. 

```{r topotag, fig.align='center',fig.width=6,fig.height = 4, warning = FALSE}
require("rerddapXtracto")
require("rerddap")
require("ggplot2")
require(mapdata)
ylim<-c(15,30)
xlim<-c(-160,-105)
topoInfo <- rerddap::info('etopo360')
topo <- rxtracto(topoInfo, parameter='altitude', xcoord=xpos, ycoord=ypos, xlen=.1, ylen=.1)
alldata <- cbind(tagData, topo)
alldata$lon<-alldata$lon-360
alldata$mean <- as.array(alldata$'mean altitude')
# get outline data for map
w <- map_data("worldHires", ylim = ylim, xlim = xlim)
z<-ggplot(alldata,aes(x=lon,y=lat)) + 
   geom_point(aes(colour=mean),size=2.) + 
   scale_shape_manual(values=c(19,1))
z + geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") + 
  theme_bw() + 
  scale_colour_gradient( name = "Depth") + 
  coord_fixed(1.3,xlim = xlim, ylim = ylim) + ggtitle("Bathymetry at marlin tag locations")

```

Again, note the differences from the `xtractomatic` example:

- the user has to find the datasetID from the given <span style="color:blue">ERDDAP</span> server, in this case 'etopo360' (from http://upwell.pfeg.noaa.gov/erddap/griddap/etopo360.html)
- the user has to call `rerddap::info('etopo360')` to obtain information about that dataset
- the user has to know the parameter name desired for that datasetID, in this case 'altitude'
- only the xcoord and ycoord values need to be given, as 'etopo360' is a 2-D dataset, and this is handled more gracefully then in `xtractomatic`.


## Using `rxtracto_3D` 

The function `rxtracto_3D()` adds no new capabilites to `rerddap`, but it does return the data in a structure consistent with the `xtractomatic` function `xtracto_3D()`, is used in the function `rxtractogon()`, and provides a consistent parallel to the functions in the `xtractomatic` package. It also changes latitudes and longitudes to agree with those of the source dataset, and returns a structure where these are mapped back to the request.

### Obtaining VIIRS chlorophyll data


```{r VIIRSchla}
require("rerddapXtracto")
require("rerddap")
#require("lubridate")
xpos<-c(235,240)
ypos<-c(36,39)
tpos<-c("2012-01-15","last")
VIIRSInfo <- rerddap::info('erdVH2chlamday')
VIIRS <- rxtracto_3D(VIIRSInfo, parameter='chla', xcoord=xpos, ycoord=ypos, tcoord=tpos)
VIIRS$time <- as.Date(VIIRS$time)
```

`rxtracto_3d()` returns a list of the form:

 * VIIRS$chla       : num [xcoord, ycoord, tcoord] 
 * VIIRS$datasetname: character string of the ERDDAP dataset ID
 * VIIRS$longitude  : num [xcoord] longitude of extact
 * VIIRS$latitude   : num [ycoord] latitudes of extract
 * VIIRS$altitude   : num zcoord of extract
 * VIIRS$time       : POSIXlt[tcoord] times of extract

The coordinate names of the structure are based on the names given in the `rxtracto_3d()` call, so may differ between datasets. A similar call to `rerddap::griddap()` will either return the data "pre-melted" (that is long-form) or only get the netcdf file and have the user read in the data.  There are tradeoffs to having the data "pre-melted", for consistency we maintain a structure similar to that in  `xtracto_3D()`, plus this works even if the xcoord and ycoord are not longitude and latitude.

We can map the results using a "helper" function that "melts" the data into "longform."


```{r mapFrame}
mapFrame<- function(longitude,latitude,chla){
  dims<-dim(chla)
  chla<-array(chla,dims[1]*dims[2])
#  longitude[longitude > 180] <- longitude[longitude > 180] - 360
  chlaFrame<-expand.grid(x=longitude,y=latitude)
  chlaFrame$chla <- chla
  return(chlaFrame)
}
```

and also define a helper function `plotFrame` to plot the data:

```{r plotFrame}
plotFrame<-function(chlaFrame, xlim, ylim,title,logplot=TRUE){
  require(mapdata)
  require(ggplot2)
  w <- map_data("worldHires", ylim = ylim, xlim = xlim)
  myplot<-ggplot(data = chlaFrame, aes(x = x, y = y, fill = chla)) +
    geom_raster(interpolate = FALSE) +
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim)
  if(logplot){
    my.col <- colors$chlorophyll
    myplot<-myplot + scale_fill_gradientn(colours = my.col, na.value = NA,limits=c(-2,4)) +
      ggtitle(title)
    }else{
     myplot<-myplot + scale_fill_gradientn(colours = my.col, na.value = NA) +
       ggtitle(title)
  }
  return(myplot)
}
```

We examine <span style="color:blue">chlorophyll</span> in June 2012:


```{r VIIRSLogPlot, fig.width=5,fig.height = 5, fig.align='center', warning=FALSE}
xlim<-c(235,240)-360
ylim<-c(36,39)
ttext<-VIIRS$time[month(VIIRS$time) == 6 & year(VIIRS$time) == 2012]
chlalogFrame<-mapFrame(VIIRS$longitude-360,VIIRS$latitude,
                       log(VIIRS$chla[,,month(VIIRS$time) == 6 & year(VIIRS$time) == 2012]))
chlalogPlot<-plotFrame(chlalogFrame,xlim,ylim,paste("VIIRS log(chla)",ttext))
chlalogPlot
```






## Using `rxtractogon` 

The `rxtractogon` function extracts a time-series of satellite data that are within a user supplied polygon. We repeat here two of the examples in the `xtractomatic` vignette.  The first gets chlorophyll within the boundary points of the  Monterey Bay National Marine Sanctuary, which are available in the `mbnms` dataset which are loaded with the `rerddapXtracto` package.  

```{r mbnmsChla}
#require(rerddapXtracto)
dataInfo <- rerddap::info('erdVH2chlamday')
parameter = 'chla'
tpos<-c("2014-09-01","2014-10-01")
#tpos <-as.Date(tpos)
xpos<-mbnms$Longitude
ypos<-mbnms$Latitude
sanctchl <- rxtractogon (dataInfo, parameter=parameter, xcoord=xpos, ycoord=ypos,  tcoord=tpos)
str(sanctchl)
```

The extract (see`str(sanctchl)`) contains two time periods of <span style="color:blue">chlorophyll</span> masked for data only in the sanctuary boundaries.  A plot of the the second time period: 

```{r mbnmsChlaPlot, fig.width=5,fig.height = 5, fig.align='center', warning = FALSE}
require(ggplot2)
require(mapdata)
xlim = c(-123.5,-121.)
ylim = c(35,38)
mbnmsFrame<-mapFrame(sanctchl$longitude,sanctchl$latitude,log(sanctchl$chla[,,2]))
my.col <- colors$chlorophyll
w <- map_data("worldHires", ylim = ylim, xlim = xlim)
myplot<-ggplot() + geom_path(data=mbnms,aes(x=Longitude,y=Latitude),colour="black")   
myplot<-myplot + 
    geom_raster(data = mbnmsFrame, aes(x = x, y = y, fill = chla),interpolate = FALSE) + 
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + scale_fill_gradientn(colours = my.col, na.value = NA,limits=c(-1,3)) +
    ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim) + 
    ggtitle(paste("log(Chla) in MBNMS",sanctchl$time[2]))
myplot
```


The MBNMS is famous for containing the Monterey Canyon, which reaches depths of up to 3,600 m (11,800 ft) below surface level at its deepest. `xtractogon` can extract the bathymetry data for the MBNMS from the <span style="color:red">ETOPO</span> dataset:

```{r mbnmsBathy}
#require(xtractomatic)
dataInfo <- rerddap::info('etopo180')
xpos<-mbnms$Longitude
ypos<-mbnms$Latitude
bathy <- rxtractogon(dataInfo, parameter='altitude', xcoord=xpos, ycoord=ypos)
str(bathy)
```

Mapping the data to show the canyon:

```{r mbnmsBathyPlot, fig.width=5,fig.height = 5, fig.align='center', warning=FALSE}
require(ggplot2)
require(mapdata)
xlim = c(-123.5,-121.)
ylim = c(35,38)
mbnmsFrame<-mapFrame(bathy$longitude,bathy$latitude,bathy$depth[,,1])
w <- map_data("worldHires", ylim = ylim, xlim = xlim)
myplot<-ggplot() + geom_path(data=mbnms,aes(x=Longitude,y=Latitude),colour="black")   
myplot<-myplot + geom_raster(data = mbnmsFrame, aes(x = x, y = y, fill = chla),interpolate = FALSE) +
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + scale_fill_gradient(na.value = NA, name ="Depth") +
    ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim) + ggtitle("MBNMS Bathymetry")
myplot
```


## Temperature at 70m in the north Pacific from the SODA model output

This is an example of an extract from a 4-D dataset (results from the "Simple Ocean Data Assimilation (SODA)" model), and illustrate the case where the z-coordinate does not have the default name "altitude".  Water temperature at 70m depth is extracted for the North Pacific Ocean.


```{r soda70}
#require(xtractomatic)
dataInfo <- rerddap::info('hawaii_d90f_20ee_c4cb')
xpos <- c(135.25, 240.25)
ypos <- c(20.25, 60.25)
zpos <- c(70.02, 70.02)
tpos <- c('2010-12-15', '2010-12-15')
soda70 <- rxtracto_3D(dataInfo, parameter='temp', xcoord=xpos, ycoord=ypos, tcoord=tpos, zcoord=zpos, zName='depth')
str(soda70)
```

SInce the data cross the dateline, it is necessary to use the new "world2Hires" in the package "mapdata" which is Pacific Ocean centered.  Unfortunatley there is a smal problem where the outlines from certain countries wrap and mistakenly appear in plots, and those countries must be removed,  see code below.


```{r soda70Plot, fig.width=6,fig.height = 3, fig.align='center', warning = FALSE}
require(ggplot2)
require(mapdata)
xlim <- c(135,240)
ylim <- c(20,60)
soda70Frame<-mapFrame(soda70$longitude,soda70$latitude,soda70$temp[,,1,1])
my.col <- colors$temperature
## Must do a kludge to remove countries that wrap and mess up the plot
w1 <- map("world2Hires", xlim=c(135, 240), ylim=c(20, 60), fill=TRUE, plot=FALSE)
remove <- c("UK:Great Britain", "France", "Spain", "Algeria", "Mali", "Burkina Faso", "Ghana", "Togo")
w <- map_data("world2Hires", regions=w1$names[!(w1$names %in% remove)], ylim = ylim, xlim = xlim)
myplot<-ggplot() + 
    geom_raster(data = soda70Frame, aes(x = x, y = y, fill = chla),interpolate = FALSE) + 
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + scale_fill_gradientn(colours = my.col, na.value = NA, limits=c(-3,30), name="temperature") +
    ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim) + 
    ggtitle(paste("temperature at 70 meters depth from SODA for",soda70$time[1]))
myplot
```


## Using a different ERDDAP server 

### Irish Marine Institute {#hourly}

The Irish Marine Institute has an ERDDAP server at http://erddap.marine.ie/erddap.  Among other datasets, they have hourly output from a model of the North Altantic ocean, with a variety of ocean related parameters, see http://erddap.marine.ie/erddap/griddap/IMI_NEATL.html.  To obtain sea surface salinity at 2016-11-03 for the domain of the model:

```{r NAtlSSS}
urlBase <- "http://erddap.marine.ie/erddap/"
parameter <- "sea_surface_salinity"
sssTimes <- c("2016-11-02T23:00:00Z","2016-11-02T23:00:00Z")
sssLats <- c(48.00625, 57.50625)
sssLons <- c(-17.99375, -1.00625)
dataInfo <- rerddap::info("IMI_NEATL",url=urlBase)
NAtlSSS <- rxtracto_3D(dataInfo, parameter=parameter, xcoord=sssLons, ycoord=sssLats, tcoord=sssTimes, urlbase=urlBase)
str(NAtlSSS)
```

```{r NAtlSSSplot, fig.width=6,fig.height = 3, fig.align='center', warning = FALSE}
require(ggplot2)
require(mapdata)
xlim <- c(-17.99375, -1.00625)
ylim <- c(48.00625, 57.50625)
NAtlSSSFrame<-mapFrame(NAtlSSS$longitude,NAtlSSS$latitude,NAtlSSS$sea_surface_salinity[,,1])
my.col <- colors$salinity
w <- map_data("worldHires", ylim = ylim, xlim = xlim)
myplot<-ggplot() + 
    geom_raster(data = NAtlSSSFrame, aes(x = x, y = y, fill = chla),interpolate = FALSE) + 
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + scale_fill_gradientn(colours = my.col, na.value = NA, limits=c(34,36), name="salinity") +
    ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim) + 
    ggtitle(paste("salinity", NAtlSSS$time[1]))
myplot
```

### IFREMER

The French agency IFREMER also has an <span style="color:blue">ERDDAP</span> server. We obtain salinity data at 75 meters from Global Ocean, Coriolis Observation Re-Analysis CORA4.1 model off the west coast of the United States.

```{r IFREMER}
urlBase <- "http://www.ifremer.fr/erddap/"
parameter <- "PSAL"
ifrTimes <- c("2013-05-15","2013-05-15")
ifrLats <- c(30., 50.)
ifrLons <- c(-140., -110.)
ifrDepth <- 75.
dataInfo <- rerddap::info("ifremer_tds0_6080_109e_ed80",url=urlBase)
ifrPSAL <- rxtracto_3D(dataInfo, parameter=parameter, xcoord=ifrLons, ycoord=ifrLats, tcoord=ifrTimes, zcoord=ifrDepth, zName='depth', urlbase=urlBase)
str(ifrPSAL)
```

The `ggplot2` function `geom_raster()` is not designed for unevenly spaced coordinates, as are the latitudes from this model.  The function `interp()` from the package `akima` is used to interpolate the data which are then plotted.


```{r ifrPSALplot, fig.width=6, fig.height = 3, fig.align='center', warning = FALSE}
## ggplot2 has trouble with unequal y's
 require(mapdata)
 require(ggplot2)
 require(akima)
 require(dplyr)
  xlim <- c(-140, -110)
  ylim <- c(30, 51)
## ggplot2 has trouble with unequal y's
  my.col <- colors$temperature
  tempData1 <- ifrPSAL$PSAL[,,1,1]
  tempData<-array(tempData1 ,61*54)
  tempFrame<-expand.grid(x=ifrPSAL$longitude,y=ifrPSAL$latitude)
  tempFrame$temp <- tempData
  tempFrame1 <- dplyr::filter(tempFrame, !is.nan(temp))
  myinterp <- akima::interp(tempFrame1$x, tempFrame1$y, tempFrame1$temp,xo=seq(min(tempFrame1$x), max(tempFrame1$x), length = 61),
       yo=seq(min(tempFrame1$y), max(tempFrame1$y), length = 54))
  myinterp1 <- expand.grid(x=myinterp$x,y=myinterp$y)
  myinterp1$temp <- array(myinterp$z, 61*54)
  w <- map_data("worldHires", ylim = ylim, xlim = xlim)
 myplot<-ggplot() +
    geom_raster(data = myinterp1, aes(x = x, y = y, fill = temp), interpolate = FALSE) +
    geom_polygon(data = w, aes(x=long, y = lat, group = group), fill = "grey80") +
    theme_bw() + scale_fill_gradientn(colours = my.col, na.value = NA, limits=c(32,35), name="salinity") +
    ylab("latitude") + xlab("longitude") +
    coord_fixed(1.3,xlim = xlim, ylim = ylim) + ggtitle(paste("salinity at 75 meters",ifrPSAL$time[1] ))
 myplot
```


## What happens when you request an extract

When you make an `rerddapXtracto` request, particularly for track data using the `rxtracto` function, it is important to understand what is extracted, because the remote dataset requested likely will have a different temporal and spatial resolution then the local dataset.

Specifically, let `longitude`, `latitude` and `time` be the coordinate system of the remote <span style="color:blue">ERDDAP</span> dataset, and let `xpos`, `ypos` and `tpos` be the bounds of a request.  Then the <span style="color:blue">ERDDAP</span> request is based on the nearest grid point of the <span style="color:blue">ERDDAP</span> dataset:

```{r nearGrid,eval=FALSE}
latitude[which.min(abs(latitude - ypos[1]))]  # minimum latitude
latitude[which.min(abs(latitude - ypos[2]))]  # maximum latitude
longitude[which.min(abs(longitude- xpos[1]))] # minimum longitude
longitude[which.min(abs(longitude - xpos[2]))] # maximum longitude
isotime[which.min(abs(time- tpos[1]))] # minimum time
isotime[which.min(abs(time - tpos[2]))] # maximum time
```

where  `tpos` and `time` have been converted to an <span style="color:blue">R</span> date format so that it is a number rather than a string. For example, the FNMOC 6-hourly Ekman transports are on a 1-degree grid. A request for the data at a longitude of 220.2 and a latitude of 38.7 will return the result at a longtiude of 220 and a latitude of 39.  These values are then used in a call to `rerddap::griddap()`, with `parameter` being the variable of interest.

